---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# hydrofabric

<!-- badges: start -->

[![R-CMD-check](https://github.com/NOAA-OWP/hydrofabric/workflows/R-CMD-check/badge.svg)](https://github.com/NOAA-OWP/hydrofabric/actions)
<!-- badges: end -->

### Overview

There are three major types of network refactoring needed to meet a broad set of needs:

  1. One based on a flowline length criteria (routing)
  2. One that aims for a uniform catchment size (rainfall-runoff)
  3. A POI version that forces things down to a set critical locations (PRMS).
  
`hydrofabric` is a set of packages that work in harmony to meet these needs. The package is designed to make it easy to install and load core packages across users and organizations in a single command.

## Installation

```{r eval = FALSE}
# Install the development version from GitHub
# install.packages("remotes")
remotes::install_github("NOAA-OWP/hydrofabric")
```

## Usage

`library(hydrofabric)` will load the core packages: 

* [nhdarrow]() (In development)
* [nwmdata]() (In development)
* [zonal](https://github.com/mikejohnson51/zonal) for catchment parameter estimation
* [nhdplusTools](https://github.com/usgs-r/nhdplusTools/) for network manipulation
* [hyRefactor](https://github.com/dblodgett-usgs/hyRefactor) for network factoring
* [hyAggregate](https://github.com/mikejohnson51/hyAggregate) for network aggregation
* [hyRelease](https://github.com/mikejohnson51/hyRelease) for data releases running elected subroutines


And you can check that all tidyverse packages are up-to-date with `tidyverse_update()`:

```{r update, eval = FALSE}

```

## Code of Conduct

Please note that the project is released with a [Contributor Code of Conduct](). 
By contributing to this project, you agree to abide by its terms.
