---
title: "The NextGen Hydrofabic Data Model"
description: |
  "What actually comes with a hydrofabric?
author:
  - name: "Mike Johnson"
    url: https://github.com/mikejohnson51
    affiliation: Lynker, NOAA-Affiliate
    affiliation_url: https://lynker.com
output: distill::distill_article
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  message = FALSE, out.width = "100%")

library(dm)
library(DT)
library(mapview)
library(hydrofabric)
library(ggplot2)
library(dataRetrieval)

dm   = hydrofabric::hf_dm$dm
meta = hydrofabric::hf_dm$meta
```

# Data Model

A data model has been developed to provide the minimal, yet efficient, types of data needed for modeling, hydrofabric subsetting, cross walking, and indexing.

Much of this design stems from the OGC report [here](https://docs.ogc.org/per/22-040.html) with modifications made for the NextGen modelling application.

The current data model contains 7 data layers with an anticipated 8th for water bodies.

5 of these are spatial and 2 are a-spatial highlighting the value of the geopackage model:

```{r, echo = FALSE, fig.height=8, fig.width=8, fig.show='hold'}
dm_draw(dm,  view_type = "all", column_types = TRUE, rankdir = "LR")
```


# 1. Divides

- [Divides](https://docs.opengeospatial.org/is/14-111r6/14-111r6.html#_catchment_divide) represent the incremental area draining to a flowpath.

- They are a one-dimensional, hydrology-specific realization of the holistic catchment. 

- Geometrically, each divide is a an edge bounded by inflow and outflow nodes forming a POLYGON


```{r, echo = FALSE}
poudre = "cihro-data/poudre.gpkg"
divides = read_sf(poudre, "divides")

divides = rename_geometry(divides, "geometry")
df = filter(hydrofabric::hf_dm$meta, hydrofabric::hf_dm$meta$Attribute %in% names(divides))

df[match(names(divides), df$Attribute),] %>% 
  datatable(options = list(pageLength = 35, dom = 't'), rownames = FALSE)
```

## Divide Type

Divides provide a continuous coverage of the modeling domain. While the majority of the domain is dendritic, there are areas where there is a divide with no flow, or, flow with no divide.

For example:
 
 - a coastal divide has an area, but no flowpath.

 - an internally draining region (sink) has an area, but no flow path

 - An artificial path (canal) has a flowpath, but no associated drainage area

> When both a divide and flowpath representation exisit, there is a `1`:`1` relationship meaning there is one `flowpath` representation for each `divide`. In this case the numeric ID of each is the same

In the future, we will support the ability to have `1:many` `divide`:`*flowline*` representations

```{r}
divides = read_sf(poudre, "divides")
mapview(divides[5,])
```

# 2. Flowpaths

 - A [flowpath](https://docs.opengeospatial.org/is/14-111r6/14-111r6.html#_flowpath_also_flow_path) references the path of a moving particle of water
 - The flowpath can be represented as an edge bounded by inflow and outflow nodes, and associated with left-bank and right-bank sub-catchment faces.
 - Geometrically, each flowpath is a LINESTRING connecting the inflow to the outflow with some level of sinuosity
 - The representation of a flowpath can change per modeling application from straight line to "infinity" sinuous
 
```{r, echo = FALSE}
flowpaths = read_sf(poudre, "flowpaths") %>% 
  rename_geometry("geometry")
df = filter(meta, meta$Attribute %in% names(flowpaths))

df[match(names(flowpaths), df$Attribute),] %>% 
  DT::datatable(options = list(pageLength = 35, dom = 't'), rownames = FALSE)
```

```{r}
# Read in all flowpaths
flowpaths = read_sf(poudre, "flowpaths")

# Define outflow
outflow  =  slice_max(flowpaths, hydroseq) 

mapview(outflow) + 
  filter(divides, divide_id == outflow$divide_id)
```

## Durable Integration

 - The majority of fields in the hydrofabric are not designed to remain resilient through the refactor/aggregation process. 
 - Some that are related to network indexing and integration are persistent though.
 - A primary one of these is the mainstem identifier
 - This data model is described in detail [here](https://www.sciencedirect.com/science/article/pii/S1364815220309841) and a video is [here](https://www.usgs.gov/media/videos/waterml2-part-3-surface-hydrology-features-hyfeatures-conceptual-model-and-mainstems)


```{r}
# Mainstem Persistent Identifier
(pid = glue("https://geoconnex.us/ref/mainstems/{outflow$mainstem}"))

# Geonconnex PID mainstem read for URL
geoconnex_mainstem = read_sf(pid)

# Mainstem defined by membership of the outlets
hf_mainstem = filter(flowpaths, mainstem == outflow$mainstem)

# View
mapview(geoconnex_mainstem) + 
  mapview(hf_mainstem, color = "red")
```
- Anything that is mapped to a shared mainstem ID (observations, models, etc) can be shared across realizations and become more "durable"

# 3. Nexus

A [nexus](https://docs.opengeospatial.org/is/14-111r6/14-111r6.html#_hydro_logic_nexus) provides a conceptual outlet for water contained by a catchment. 

They are the locations where NextGen will exchange information!

  - some are defined by POIs
  
  - Others are defined by locations in a network where a 1:1 inflow and outflow occur

Every catchment flows to a nexus, conversely every location in a hydrologic system can be thought of as a hydro nexus that drains some catchment(s).

```{r, echo = FALSE}
nexus = read_sf(poudre, "nexus")
nexus = rename_geometry(nexus, "geometry")
df = filter(meta, meta$Attribute %in% names(nexus))

df[match(names(nexus), df$Attribute),] %>% 
  DT::datatable(options = list(pageLength = 35, dom = 't'), rownames = FALSE)
```

## Consequense of Network Aggregation

More then one divide/flowpath (e.g. catchment) can contribute to a single nexus:

```{r, echo = FALSE}
count = sort(table(divides$toid), decreasing= TRUE) 
```

```{r, echo = FALSE}
ggplot() +
  geom_bar(aes(x = 100* as.vector(count) /length(count))) + 
  labs(y = "Percert of Nexus(s)", x = "Number of contributing catchments") + 
  theme_light()
```


```{r}
count = sort(table(divides$toid), decreasing= TRUE) 

(example = names(count[count == 3][1]))

(sub_divides = filter(divides, toid == example))
```

```{r, echo = FALSE}
mapview(sub_divides) + 
  filter(nexus, id == example) + 
  filter(flowpaths, toid == example)
```


# 4. Hydrolocations

- A [H=hydrolocations](https://docs.opengeospatial.org/is/14-111r6/14-111r6.html#_hydro_logic_location) can define any location of hydrologic significance located "on" a hydrologic network

- In a given dataset, hydrolocations may or may not have an associated hydrologic nexus and catchment feature. 

 - In such cases, hydro locations would typically be linearly-referenced to a defined set of flowpaths. 

- Topologically, a hydro-location can be understood as an inlet or outlet node located at the end of a flowpath edge.

- In NextGen design, a subset of the community POIs, are treated as hydrolocations in which the network CANNOT be refactored or aggregated "over".


```{r, echo = FALSE}
hydrolocations = read_sf(poudre, "hydrolocations") %>% 
  rename_geometry( "geometry")
df = filter(meta, meta$Attribute %in% names(hydrolocations))

df[match(names(hydrolocations), df$Attribute),] %>% 
  DT::datatable(options = list(pageLength = 35, dom = 't'), rownames = FALSE)
```

```{r}
hl = filter(hydrolocations, id == outflow$toid)

filter(hydrolocations, hl_uri == "Gages-06752260")
glimpse(hl)

(pid = glue('https://reference.geoconnex.us/collections/{tolower(hl$hl_reference)}/items?provider_id={hl$hl_link}'))

pid = read_sf(pid)

glimpse(pid)
```

## Multiplicity

The data design of the hydrolocation location layer intends to represent each realization of a POI independently.

This means we can have a shared `hl_id` and `id` (nexus location) but distinct `hl_reference`, and `hl_link`

```{r, echo = FALSE}
table = table(hydrolocations$hl_id) %>% 
  sort(decreasing = TRUE) 
```

```{r}
(poi = filter(hydrolocations, hl_id == 7253))
```

## Rigorous community POI mapping 

The geoconnex PID can provide the location of the source data.

Here we can see the community POI was moved slightly to integrate with the reference network:


```{r}
geoconnex = read_sf('https://reference.geoconnex.us/collections/dams/items?provider_id=CO01659')

mapview(geoconnex, color = "red") + poi
```

# 5. Network

The network layer is tabular and has no spatial information

It provides the ultimate cross walk between the NextGen fabric, the source hydrofabric, and all mapped hydrolocations.

```{r, echo = FALSE}
network = read_sf(poudre, "network") 
df = filter(meta, meta$Attribute %in% names(network))

df[match(names(network), df$Attribute),] %>% 
  DT::datatable(options = list(pageLength = 35, dom = 't'), rownames = FALSE)
```

This network is key to hydroindexing, data query, and network subsetting. 

### Example:

_What is the NextGen ID near me?_

Lets say we lived at this location in Laporte, CO

```{r}
pt = st_sfc(st_point(c(-105.14044,  40.62949)), crs = 4326)
mapview(pt)
```

We can use the NLDI to map the reference fabric to this point, and then search for how the reference fabric hf_id merged into the current network:

```{r}
(x = findNLDI(location = pt))

(n = filter(read_sf(poudre, "network"), hf_id == x$origin$comid))

mapview(filter(flowpaths, id == n$id)) + 
  filter(divides, divide_id == n$divide_id) +
  filter(nexus, id == n$toid) + 
  pt
```

# 6. Flowpath Attributes

- Flowpath attributes are extracted from the NWM Routelink file.

- The network layer is used to "length weight average" these attributes to the new network.

- Ongoing advances in modeling the 3D hydrofabric network will soon eclipse this approach.


# 7. Lakes

Lake attributes are extracted from the NWM Lake file file.

The network layer is used to "length weight average" these attributes to the new network.



# 8. Layer Style

- When you open a hydrofabric gpkg created with many of the `hydrofab` tools, you will find they load with a consistent symbology.

- These symbologies are saved with the Geopackage using the `hydrofab::append_style()`

- Built on reference [QGIS QML files](https://github.com/mikejohnson51/hydrofab/tree/master/inst/qml).

 