---
title: "The NextGen Hydrofabic Data Model"
description: |
  "What actually comes with a hydrofabric?
author:
  - name: "Mike Johnson"
    url: https://github.com/mikejohnson51
    affiliation: Lynker, NOAA-Affiliate
    affiliation_url: https://lynker.com
output: distill::distill_article
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  message = FALSE, out.width = "100%")

library(dm)
library(DT)
library(mapview)
library(hydrofabric)
library(ggplot2)
library(dataRetrieval)

hf_dm = list(
  
  flowlines = data.frame(
    id = integer(1L),
    toid = integer(1L),
    mainstem = integer(1L),
    order = integer(1L),
    hydroseq = integer(1L),
    lengthkm = numeric(1L),
    areasqkm = numeric(1L),
    tot_drainage_areasqkm = numeric(1L),
    divide_id = integer(1L),
    has_divide = logical(1),
    geometry = double(1L)
  ),
  
  flowpath_attributes = data.frame(
    id = integer(1L),
    rl_gages = character(1L),
    NHDWaterbodyComid = character(1L),
    Qi = numeric(1L),
    MusK = numeric(1L),
    MusX = numeric(1L),
    n = numeric(1L),
    So = numeric(1L),
    ChSlp = numeric(1L),
    BtmWdth = numeric(1L),
    Kchan = numeric(1L),
    nCC = numeric(1L),
    TopWdthCC = numeric(1L),
    TopWdth = numeric(1L),
    length_m = numeric(1L)
  ),
  

  lakes = data.frame(
    id           = integer(1L),    
    toid  = integer(1L),
    hl_id = integer(1L),
    hl_reference = character(1L),
    hl_link      = character(1L),
    hl_uri       = character(1L),
    Dam_Length  = numeric(1L) ,
    ifd  = numeric(1L),
    LkArea  = numeric(1L),
    LkMxE  = numeric(1L),
    OrificeA  = numeric(1L),
    OrificeC  = numeric(1L),
    OrificeE   = numeric(1L),
    time = numeric(1L),
    WeirC = numeric(1L), 
    WeirE = numeric(1L), 
    WeirL = numeric(1L), 
    geometry = double(1L)
  ),
  
  
  divides = data.frame(
    divide_id = integer(1L),
    id   = integer(1L),
    toid = integer(1L),
    ds_id = integer(1L),
    areasqkm = numeric(1L),
    type = character(1L),
    has_flowline = logical(1),
    geometry = double(1L)
  ),
  
  nexus = data.frame(
    id      = integer(1L),
    toid    = integer(1L),
    hl_id   = character(1L),
    hl_uri      = character(1L),
    type    = character(1L),
    geometry = double(1L)
  ),
  
  # Can have 1:many, NOT! part of topology 
  hydrolocations = data.frame(
    hl_id = integer(1L),
    id           = integer(1L),
    hl_reference = character(1L),
    hl_link      = character(1L),
    hl_uri       = character(1L),
    hl_position  = character(1L),
    geometry     = numeric(1L)
  ),
  
  network = data.frame(
    # Topology
    id         = integer(1L),
    toid       = integer(1L),
    # Associations
    divide_id  = integer(1L),
    ds_id = integer(1L),
    mainstem   = integer(1L),
    hydroseq   = integer(1L),
    wb_id      = integer(1L), 
    hl_id      = integer(1L),
    hl_uri     = character(1L),
    # Reference
    hf_source  = integer(1),
    hf_id      = numeric(1L),
    #hf_id_part = integer(1L),
    # Description
    lengthkm = numeric(1L),
    areasqkm = numeric(1L),
    tot_drainage_areasqkm = numeric(1L),
    type = character(1L),
    vpu = character(1L)
  ),

  WB = data.frame(
    id    = integer(1L),
    wb_id = integer(1L),
    wb_area = numeric(1L),
    wb_source = numeric(1L),
    geometry = numeric(1L)
  ) 
)


meta = tibble::tribble(
~Attribute, ~Description,
"id",             "A hydrofabric specfic, globaly unique flowline identifier",
"hf_source",      "The source dataset for the hydrofabric development (e.g. NHDPlusV2)",
"hf_id",          "The unique identifier in the source hydrofabric (hf_source)",
#"hf_id_part",     "If the original hydrofabric identifier was split, the sub part. Parts increase from outlet to inlet",
"divide_id",      "A hydrofabric specfic, globaly unique divide identifier",
"mainstem",       "A nationally unique identifier assigned to the set of flowlines that compose a stream from its headwater to its mouth",
"hl_id",          "A hydrofabric specifc, globaly unique, hydrologic location identifier",
"hl_reference",   "The hydrologic location type (e.g. Gage, NID)",
"hl_link",         "The unique identifier in the source hydrologic location dataset.",
"hl_uri",         "A comma seperated conncatination of hl_reference and hl_link that mirrors thse used in the NLDI",
"hl_position",    "Position of hydrolocation on flowpath (inflow, outflow, along)",
"wb_id",          "Water body Common Identifier from wb_source",
"ds_id",          "Most downstream adjacent divide. Only applicable to internal catchments",

"toid",                     "The identifier of the directly downstream flowpath/flowline",
"lengthkm",                 "The length in kilometers of the flowpath element",
"areasqkm",                 "The area of the incremental divides",
"tot_drainage_areasqkm",    "The total upstream area contributing to the feature",

"order",                    "Strahler stream order",
"hydroseq",                 "VPU based hydrologic sort. Increases from downstream to upstream",

"geometry",      "Simple Features Geometry (POINT, LINESTRING, POLYGON)",
"type",          "Type of network feature (network, internal, coastal, connector)",

"wb_area",       "Waterbody area",
"wb_source",     "Waterbody source",

"has_divide",    "Does an abstract catchment have a divide realization",
"has_flowline",  "Does an abstract catchment have a flowline/flowpath realization",
"vpu", "A processing unit used to segment large scale hydrofabrics",


    'rl_gages' , 'NHD Gage Event ID from SOURCE_FEA field in Gages feature class',
    'NHDWaterbodyComid' , 'ComID of NHDWaterbody feature associated using spatial join (intersection) between NHDFlowline_network and Waterbodies"',
    'Qi' , 'Length weighted Initial flow in link (CMS)',
    'MusK' , 'Length weighted  Muskingum routing time (s)',
    'MusX' , 'Length weighted  Muskingum weighting coefficient',
    'n' ,"Length weighted Manning's roughness" ,
    'So' , "Slope computed from the aggregated flow network",
    'ChSlp' ,"Length weighted Channel side slope" ,
    'BtmWdth' ,"Length weighted Bottom width of channel (m)" ,
    'Kchan' , "Length weighted channel conductivity",
    'nCC' , "Length weighted Compound Channel Manning's n",
    'TopWdthCC' , "Compound Channel Top Width (m)",
    'TopWdth' , "Length weighted Top Width (m)",
    'length_m' , "Length computed from the aggregated flow network",

    'Dam_Length'  , "Dam Length (m)" ,
    'ifd'  , "Initial fraction water depth",
    'LkArea'  , "Lake area (sq. km)",
    'LkMxE'  , "Maximum lake elevation (m ASL)",
    'OrificeA'  , "Orifice cross-sectional area (sq. m)",
    'OrificeC'  , "Orifice coefficient",
    'OrificeE'   , "Orifice elevation (m ASL)",
    'time' , "time of measurement",
    'WeirC' ,"Weir coefficient" , 
    'WeirE' ,"Weir elevation (m ASL)" , 
    'WeirL' , "Weir length (m)")

x = unique(unlist(sapply(hf_dm, names)))


if(!all(length(x[!x %in% meta$Attribute]) == 0, length(meta$Attribute[!meta$Attribute %in% x]) == 0)){stop()}

dm = dm::dm(
            network = hf_dm$network,
            flowlines = hf_dm$flowlines,
            hydrolocations = hf_dm$hydrolocations,
            divides = hf_dm$divides,
            WB = hf_dm$WB,
            lakes = hf_dm$lakes,
            nexus = hf_dm$nexus,
            flowpath_attributes = hf_dm$flowpath_attributes) %>% 
  dm_add_pk(flowlines, id)  %>% 
  dm_add_pk(hydrolocations, hl_id)  %>% 
  dm_add_pk(WB, wb_id)  %>% 
  dm_add_pk(divides, divide_id)  %>% 
  dm_add_pk(nexus, id)  %>% 
  dm_add_pk(flowpath_attributes, id)  %>% 
  dm_add_pk(lakes, id)  %>% 
  dm_set_colors(red = flowlines, 
                red = divides,
                red = hydrolocations,
                gray = WB, 
                red  = network,
                blue = flowpath_attributes,
                blue = nexus, 
                blue = lakes)

```


# Data Model

A rigorous data model has been developed to provide the minimal, yet efficient, types of data needed for modeling, hydrofabric subsetting, cross walking, and indexing.

Much of this design stems from the OGC report [here](https://docs.ogc.org/per/22-040.html) with modifications made for the NextGen modelling application.

The current data model contains 7 data layers with an anticipated 8th for water bodies.

5 of these are spatial and 2 are a-spatial highlighting the value of the geopackage model:

```{r, echo = FALSE, fig.height=8, fig.width=8, fig.show='hold'}
dm_draw(dm,  view_type = "all", column_types = TRUE, rankdir = "LR")
```


# 1. [Divides](https://docs.opengeospatial.org/is/14-111r6/14-111r6.html#_catchment_divide)

- Divides represent the incremental area draining to a flowpath.

- They are a one-dimensional, hydrology-specific realization of the holistic catchment. 

- Geometrically, each divide is a an edge bounded by inflow and outflow nodes forming a POLYGON


```{r, echo = FALSE}
poudre = "cihro-data/poudre.gpkg"
divides = read_sf(poudre, "divides")

divides = rename_geometry(divides, "geometry")
df = filter(meta, meta$Attribute %in% names(divides))

df[match(names(divides), df$Attribute),] %>% 
  datatable(options = list(pageLength = 35, dom = 't'), rownames = FALSE)
```

## Divide Type

Divides provide a continuous coverage of the modeling domain. While the majority of the domain is dendritic, there are areas where there is a divide with no flow, or, flow with no divide.

For example:
 
 - a coastal divide has an area, but no flowpath.

 - an internally draining region (sink) has an area, but no flow path

 - An artificial path (canal) has a flowpath, but no associated drainage area

> When both a divide and flowpath representation exisit, there is a `1`:`1` relationship meaning there is one `flowpath` representation for each `divide`. In this case the numeric ID of each is the same

In the future, we will support the ability to have `1:many` `divide`:`*flowline*` representations

```{r}
divides = read_sf(poudre, "divides")
mapview(divides[5,])
```

# 2. [Flowpaths](https://docs.opengeospatial.org/is/14-111r6/14-111r6.html#_flowpath_also_flow_path)

 - The flowpath references the path of a moving particle of water
 - The flowpath can be represented as an edge bounded by inflow and outflow nodes, and associated with left-bank and right-bank sub-catchment faces.
 - Geometrically, each flowpath is a LINESTRING connecting the inflow to the outflow with some level of sinuosity
 - The representation of a flowpath can change per modeling application from straight line to "infinity" sinuous
 

```{r, echo = FALSE}
flowpaths = read_sf(poudre, "flowpaths") %>% 
  rename_geometry("geometry")
df = filter(meta, meta$Attribute %in% names(flowpaths))

df[match(names(flowpaths), df$Attribute),] %>% 
  DT::datatable(options = list(pageLength = 35, dom = 't'), rownames = FALSE)
```

```{r}
# Read in all flowpaths
flowpaths = read_sf(poudre, "flowpaths")

# Define outflow
outflow  =  slice_max(flowpaths, hydroseq) 

mapview(outflow) + 
  filter(divides, divide_id == outflow$divide_id)
```

## Durable Integration

 - Many fields in the hydrofabric are designed to remain resilient through the refactor/aggregation process. 
 - Many of these are related to network indexing and integration
 - A primary one of these is the mainstem identifier that is intended to remain persistent across network manipulations
 - This data model is described in detail [here](https://www.sciencedirect.com/science/article/pii/S1364815220309841) and a video is [here](https://www.usgs.gov/media/videos/waterml2-part-3-surface-hydrology-features-hyfeatures-conceptual-model-and-mainstems)


```{r}
# Mainstem Persistent Identifier
(pid = glue("https://geoconnex.us/ref/mainstems/{outflow$mainstem}"))

# Geonconnex PID mainstem read for URL
geoconnex_mainstem = read_sf(pid)

# Mainstem defined by membership of the outlets
hf_mainstem = filter(flowpaths, mainstem == outflow$mainstem)

# View
mapview(geoconnex_mainstem) + 
  mapview(hf_mainstem, color = "red")
```

- Anything that is mapped to a shared mainstem ID (observations, models, etc) can be shared across realizations and become more "durable"

# 3. [Nexus](https://docs.opengeospatial.org/is/14-111r6/14-111r6.html#_hydro_logic_nexus)

A nexus provides a conceptual outlet for water contained by a catchment. 

They are the locations where NextGen will exchange information!

  - some are defined by POIs
  
  - Others are defined by locations in a network where a 1:1 inflow and outflow occur

Every catchment flows to a hydro nexus, conversely every location in a hydrologic system can be thought of as a hydro nexus that drains some catchment(s).

```{r, echo = FALSE}
nexus = read_sf(poudre, "nexus")
nexus = rename_geometry(nexus, "geometry")
df = filter(meta, meta$Attribute %in% names(nexus))

df[match(names(nexus), df$Attribute),] %>% 
  DT::datatable(options = list(pageLength = 35, dom = 't'), rownames = FALSE)
```

## Consequense of Network Aggregation

More then one divide/flowpath (e.g. catchment) can contribute to a single nexus:

```{r, echo = FALSE}
count = sort(table(divides$toid), decreasing= TRUE) 
```

```{r, echo = FALSE}
ggplot() +
  geom_bar(aes(x = 100* as.vector(count) /length(count))) + 
  labs(y = "Percert of Nexus(s)", x = "Number of contributing catchments") + 
  theme_light()
```


```{r}
count = sort(table(divides$toid), decreasing= TRUE) 

(example = names(count[count == 3][1]))

(sub_divides = filter(divides, toid == example))
```

```{r, echo = FALSE}
mapview(sub_divides) + 
  filter(nexus, id == example) + 
  filter(flowpaths, toid == example) + 
  mapview(slice_max(sub_divides, tot_drainage_areasqkm), color = "red", col.regions = NA)
```


# 4. [Hydrolocations](https://docs.opengeospatial.org/is/14-111r6/14-111r6.html#_hydro_logic_location)

- A hydrolocation can define any location of hydrologic significance located "on" a hydrologic network

- In a given dataset, hydrolocations may or may not have an associated hydrologic nexus and catchment feature. 

 - In such cases, hydro locations would typically be linearly-referenced to a defined set of flowpaths. 

- Topologically, a hydro-location can be understood as an inlet or outlet node located at the end of a flowpath edge.

- In NextGen design, a subset of the community POIs, are treated as hydrolocations in which the network CANNOT be refactored or aggregated "over".


```{r, echo = FALSE}
hydrolocations = read_sf(poudre, "hydrolocations") %>% 
  rename_geometry( "geometry")
df = filter(meta, meta$Attribute %in% names(hydrolocations))

df[match(names(hydrolocations), df$Attribute),] %>% 
  DT::datatable(options = list(pageLength = 35, dom = 't'), rownames = FALSE)
```

```{r}
hl = filter(hydrolocations, id == outflow$toid)

filter(hydrolocations, hl_uri == "Gages-06752260")
glimpse(hl)

(pid = glue('https://reference.geoconnex.us/collections/{tolower(hl$hl_reference)}/items?provider_id={hl$hl_link}'))

pid = read_sf(pid)

glimpse(pid)
```

Collectively can help find, and link data!

```{r}
#Use the hl_link to extract this weeks rain event
instFlow <- dataRetrieval::readNWISdata(
  sites = hl$hl_link, service = "iv",
  parameterCd = "00060",
  startDate = "2023-05-10"
)

# Use PID COMID to extract the 42 year NWM flow record
nwm21 = nwmTools::readNWMdata(comid = pid$nhdpv2_comid)
```

```{r, echo = FALSE}
ggplot() + 
  geom_line(data= instFlow, aes(x = dateTime, y = X_00060_00000)) + 
  theme_light()
```

```{r, echo = FALSE}
ggplot() + 
  geom_line(data= nwm21, aes(x = dateTime, y = flow_cms_v2.1)) + 
  theme_light()
```

## Multiplicity

The data design of the hydrolocation location layer intends to represent each realization of a POI independently.

This means we can have a shared `hl_id` and `id` (nexus location) but distinct `hl_reference`, and `hl_link`

```{r, echo = FALSE}
table = table(hydrolocations$hl_id) %>% 
  sort(decreasing = TRUE) 
```

```{r}
(poi = filter(hydrolocations, hl_id == 7253))
```

## Rigorous community POI mapping 

The geoconnex PID can provide the location of the source data.

Here we can see the community POI was moved slightly to integrate with the reference network:


```{r}
geoconnex = read_sf('https://reference.geoconnex.us/collections/dams/items?provider_id=CO01659')

mapview(geoconnex, color = "red") + poi
```

# 5. Network

The network layer is tabular and has no spatial information

It provides the ultimate cross walk between the NextGen fabric, the source hydrofabric, and all mapped hydrolocations.

```{r, echo = FALSE}
network = read_sf(poudre, "network") 
df = filter(meta, meta$Attribute %in% names(network))

df[match(names(network), df$Attribute),] %>% 
  DT::datatable(options = list(pageLength = 35, dom = 't'), rownames = FALSE)
```

This network is key to hydroindexing, data query, and network subsetting. 

### Example:

_What is the NextGen ID near me?_

Lets say we lived at this location in Laporte, CO

```{r}
pt = st_sfc(st_point(c(-105.14044,  40.62949)), crs = 4326)
mapview(pt)
```

We can use the NLDI to map the reference fabric to this point, and then search for how the reference fabric hf_id merged into the current network:

```{r}
(x = findNLDI(location = pt))

(n = filter(read_sf(poudre, "network"), hf_id == x$origin$comid))

mapview(filter(flowpaths, id == n$id)) + 
  filter(divides, divide_id == n$divide_id) +
  filter(nexus, id == n$toid) + 
  pt
```



# 6. Flowpath Attributes

- Flowpath attributes are extracted from the NWM Routelink file.

- The network layer is used to "length weight average" these attributes to the new network.

```{r, echo = FALSE}
flowpath_attributes = read_sf(poudre, "flowpath_attributes")

df = filter(meta, meta$Attribute %in% names(flowpath_attributes))

df[match(names(flowpath_attributes), df$Attribute),] %>% 
  DT::datatable(options = list(pageLength = 35, dom = 't'), rownames = FALSE)
```


# 7. Lakes

Lake attributes are extracted from the NWM Lake file file.

The network layer is used to "length weight average" these attributes to the new network.

```{r}
lakes = read_sf(poudre, "lakes") %>% 
  filter(hl_id == 7253) %>% 
  glimpse()
```


```{r, echo = FALSE}
df = filter(meta, meta$Attribute %in% names(lakes))

df[match(names(lakes), df$Attribute),] %>% 
  DT::datatable(options = list(pageLength = 35, dom = 't'), rownames = FALSE)
```

# 8. Layer Style

- When you open a hydrofabric gpkg created with many of the `hydrofab` tools, you will find they load with a consistent symbology.

- These symbologies are saved with the Geopackage using the `hydrofab::append_style()`

- Built on reference [QGIS QML files](https://github.com/mikejohnson51/hydrofab/tree/master/inst/qml).

```{r}
read_sf(poudre, "layer_styles") %>% 
glimpse()
```


```{r}
a = system.file("qml", "divides.qml", package = "hydrofab")

readLines(a)[1:20]
```
 