---
title: "Building your own subset"
description: |
  "Automated Network Indexing, Traversal, and Data Access"
author:
  - name: "Mike Johnson"
    url: https://github.com/mikejohnson51
    affiliation: Lynker, NOAA-Affiliate
    affiliation_url: https://lynker.com
  - name: "Justin Singh"
    url: https://github.com/program--
    affiliation: Lynker, NOAA-Affiliate
    affiliation_url: https://lynker.com
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(hydrofabric)
library(DBI)
library(RSQLite)
library(mapview)
```

We are now at the point where we have a common understanding of how hydrofabrics are sourced, manipulated for a given modeling task, and generated on a VPU basis.

This aim of this section is that each of us can build a subset network for a location of interest. Doing this requires an understanding of the following:

1. The CONUS network
2. Finding an Origin
3. The R based `hfsubsetR::get`
4. The CLI utility for micro service subsets.
4. RESTFUL API

# The CONUS network file

A network file is distributed with each `version`/`type` of a hydrofabric. For more on data access patterns see the [Data Vignette](file:///Users/mjohnson/github/hydrofabric/docs/articles/data.html).

```{r}
local   <- "/Users/mjohnson/hydrofabric"
s3      <- "s3://lynker-spatial/hydrofabric"
version <-  'v2.1.1'
type    <- "nextgen"
domain  <- "conus"


network_path = glue("{local}/{version}/{type}/conus_network")
net = open_dataset(network_path)
```

In the above `schema` you'll see that every relationship between the features in the current hydrofabric, the source hydrofabric, and the conus hydrolocations, have been exploded into a "many-to-many" table. 

For example, we can look at the flowpath `wb-1002` and find that it is defined by the aggregation of NHDPlusV2 COMID 1712220, 1712230, and 1712238.

```{r}
glimpse(filter(net, id == "wb-1002"))
```
Or, that the terminal outflow of 'HUC12-010100100101' occurs at `tnx-1000000569` which is fed by an aggregate flowpath made up of three source flowpaths (`hf_id={816563, 816417, 816415}`)

```{r}
glimpse(filter(net, hl_uri == 'HUC12-010100100101'))
```

# Finding an Origin

#### By known COMID
```{r}
findOrigin(network_path, comid = 101)
```

#### By known ID
```{r}
findOrigin(network_path, id = 'wb-2430837')
```


#### By location (XY)
```{r}
here = AOI::geocode("National Water Center, Alabama")

findOrigin(network_path, xy = c(here$x, here$y))
```

#### By Hydrolocation URI
```{r}
# For a gage in Calfornia
findOrigin(network_path, hl_uri = "Gages-11123000")
# For the HUC12 of Atascadero Creek in Santa Barbara:
findOrigin(network_path, hl_uri = "HUC12-180600130201")
# For the dam on Horsetooth Reservoir
findOrigin(network_path, hl_uri = "NID-CO01659-1")
```

#### By NLDI Feature
```{r}
# For a gage in Calfornia
findOrigin(network_path, 
           nldi_feature  = list(featureSource = "nwissite", 
                                featureID = "USGS-05428500"))
```

# hfsubetR

The `hfsubsetR` library is core module in the `NOAA-OWP/hydrofabric` suite. It expedites the process of (1) finding an origin (2) traversing the network and (3) extracting the relevant features from the requested.

For example, lets get a basic subset of the network upstream of comid=101 complete with divides, nexus locations and flowlines:

```{r}
subset101 = get_subset(comid = 101, 
                       lyrs = c("divides", "nexus", "flowlines"),
                       source   = "/Users/mjohnson/hydrofabric",
                       hf_version =  '2.1.1',
                       type    = "nextgen")

mapview::mapview(subset101)

```

The same request can be appended to request attribute information to enrich the network. For example, we can grab our precomputed divide attributes and the forcing weights needed for NGIAB.

```{r}
subset101_enhanced = get_subset(comid = 101, 
           lyrs = c("divides", "nexus", "flowlines", "forcing-weights", "model-attributes"),
           source   = "/Users/mjohnson/hydrofabric",
           hf_version =  '2.1.1',
           type    = "nextgen",
           domain  = "conus")

subset101_enhanced$`model-attributes`

subset101_enhanced$`forcing-weights`

```


# CLI Option

For those interested in using the NOAA NextGen fabric as is, we have provided a Go-based CLI [here](https://github.com/lynker-spatial/hfsubset/releases)

This utility has the following syntax:

```{bash, eval = FALSE}
hfsubset - Hydrofabric Subsetter

Usage:
  hfsubset [OPTIONS] identifiers...
  hfsubset (-h | --help)

Example:
  hfsubset -l divides -o ./poudre-divides.gpkg -r "pre-release" -t hl "Gages-06752260"
  hfsubset -o ./poudre-all.gpkg -t hl "Gages-06752260"

Options:
  -l string
        Layers to subset (default "divides,nexus,flowpaths,network,hydrolocations")
  -o string
        Output file name (default "hydrofabric.gpkg")
  -r string
        Hydrofabric version (default "pre-release")
  -t string
        One of: hf, hl, or comid (default "hf")
```


# REST Service (BETA)

```{r,eval = FALSE}
https://www.lynker-spatial.com/hydrofabric/hfsubset/subset?identifier=101&identifier_type=comid
```