---
title: "Cloud Native Hydrofabric Data"
author:
  - name: "Mike Johnson"
    url: https://github.com/mikejohnson51
    affiliation: Lynker, NOAA-Affiliate
    affiliation_url: https://lynker.com
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Cloud Native Hydrofabric Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

```{r, echo = F, message = FALSE, warning=FALSE}
#source("../runners/config.R")

library(leaflet)
library(glue)
library(DT)
library(dplyr)
```

```{r, echo = FALSE, message = FALSE, warning=FALSE}

current_version = 'v2.1.1'
source = "/Users/mjohnson/hydrofabric/"

hook = glue('{source}/{current_version}/nextgen')

conus_net = arrow::open_dataset(glue('{hook}/conus_network')) %>% 
  select(vpuid, id, lengthkm, areasqkm) %>% 
  distinct() %>% 
  collect()


tmp = conus_net %>% 
  group_by(vpuid) %>% 
  summarise(total_miles = units::set_units(units::set_units(sum(lengthkm, na.rm = T), "km"), "miles"),
            total_area_miles2  = units::set_units(units::set_units(sum(lengthkm, na.rm = T), "km2"), "miles2")) %>% 
  ungroup() %>% 
  filter(!is.na(vpuid))


fl = arrow::open_dataset(glue('{hook}/conus_flowlines')) %>% 
  select(vpuid, id) %>% 
  distinct() %>% 
  group_by(vpuid) %>% 
  tally(name = 'flowpaths') %>% 
  collect()

div = arrow::open_dataset(glue('{hook}/conus_divides')) %>% 
  select(vpuid, id) %>% 
  distinct() %>% 
  group_by(vpuid) %>% 
  tally(name = "divides") %>% 
  collect()

nex = arrow::open_dataset(glue('{hook}/conus_nexus')) %>% 
  select(vpuid, id) %>% 
  distinct() %>% 
  group_by(vpuid) %>% 
  tally(name = "nexus") %>% 
  collect()

hl = arrow::open_dataset(glue('{hook}/conus_hydrolocations')) %>% 
  select(vpuid, id) %>% 
  distinct() %>% 
  group_by(vpuid) %>% 
  tally(name = "hydrolocations") %>% 
  collect()

meta = powerjoin::power_full_join(list(tmp, fl, div, nex, hl)) %>% 
  filter(!is.na(vpuid))

pop <- paste(
      paste0('<strong>GPKG: </strong>', meta$vpu ),
      paste("<strong>Flowpath:</strong>", meta$flowpaths),
      paste("<strong>Divides:</strong>", meta$divides),
      paste("<strong>Nexus Locations:</strong>", meta$nexus),
      paste("<strong>Hydrolocations:</strong>", meta$hydrolocations),
      paste("<strong>Total Flow Length (mile):</strong>", round(meta$total_miles)),
      paste("<strong>Total Divide Area (mile2):</strong>", round(meta$total_area_miles2)),
      sep = "<br/>"
    )


vpus = nhdplusTools::get_boundaries() %>%
  filter(VPUID %in% meta$vpuid) %>% 
  sf::st_transform('+proj=longlat +datum=WGS84')

labels = paste("VPU", vpus$VPUID)

bbox = as.numeric(sf::st_bbox(vpus))

```

Cloud Native NextGen hydrofabric are distributed as _NHDPlusV2_ **V**ector **P**rocessing **U**nits hive partitioned (geo)parquet datasets. They are publicly available  through `lynker-spatial`. Please note the [data license](https://lynker-spatial.s3.amazonaws.com/copyright.html) of these artifacts.

Cloud-native hydrofabric artifacts are publicly available (and egress free!) through [lynker-spatial](https://www.lynker-spatial.com/) under an [ODbL](https://opendatacommons.org/licenses/odbl/summary/) license. If you use data, please ensure you (1) Attribute Lynker-Spatial, (2) keep the data open, and that (3) any works produced from this data offer that adapted database under the ODbL. 

All data are distributed as hive partitioned (geo)parquet datasets and access follows the general pattern of: 

```{r pathing, eval = FALSE}
"{source}/{version}/{type}/{domain}_{layer}"
```

Where:

* `source` is the local or s3 location
* `version` is the release number (e.g. v2.2)
* `type` is the type of fabric (e.g. reference, nextgen, etc)
* `domain` is the region of interest (e.g. conus, hawaii, alaska)
* `layer` is the layer of the hydrofabric (e.g. divides, flowlines, network, model-attributes, routelink, etc.)

> The current version of this data is 2.1.1 (v2.1.1)

### Examples

```r
# CONUS Network Parquet
s3://lynker-spatial/hydrofabric/{version}/{type}/{domain}_network


# Model Attributes
s3://lynker-spatial/hydrofabric/{version}/{type}/{domain}s_model-attributes/
```

### Syncing to Local

AWS CLI tools can be used to sync a remote s3 directory with a local archive ensuring that you local data is up to date with the remote, assuming you want to work locally.

* The current `v2.2/reference` directory is about 3.0 GB
* The current `v2.1.1/nextgen` directory is about 8.0 GB

```{r sync}
local   <- "/Users/mjohnson/hydrofabric"
s3      <- "s3://lynker-spatial/hydrofabric"
version <-  'v2.1.1'
type    <- "nextgen"
domain  <- "conus"

(sys <- glue::glue("aws s3 sync {s3}/{version}/{type} {local}/{version}/{type}"))
```

## Interactive Viewer (v2.1.1/nextgen)

```{r, echo = FALSE}
#pal <- colorNumeric("BuPu", domain  = AOI$count, n = 10)

leaflet(width='100%') %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(data = vpus, 
              fillColor  = "gray", 
              color = "navy",
              fillOpacity = .3,
              weight = 1, 
              label = labels,
              popup = pop,
              highlightOptions = highlightOptions(color = "#FEBC11", weight = 5, bringToFront = FALSE, opacity = 1)) %>% 
  setMaxBounds(lng1 = bbox[1], lng2 = bbox[3], lat1 = bbox[2], lat2 = bbox[4])


```

```{r, echo = FALSE}

getTotal <- function(index, data){

  if(index < 1 || index > ncol(data)){
    return("")
  }  
  
  col = tryCatch({
     col <- as.numeric(as.vector(data[,index])[[1]])
  }, error = function(e){
    NULL
  })
 
  
  if(is.null(col)){
    return("")
  } else {
      return(sum(col))
  }
}





m = rbind(mutate(meta, across(names(meta), as.character)),
          sapply(1:ncol(meta), function(x){ as.character(getTotal(x, meta)) }))

DT = DT::datatable(m, escape = FALSE, width="100%",  filter = "none", 
                  rownames = T,
                options = list(autoWidth = T, 
                              pageLength = 25, 
                              scrollCollapse = T,
                              dom = 'lftp',
                              columnDefs = list(list(visible = F, targets = 0)))) %>% 
  formatCurrency('flowpaths',currency = "", interval = 3, mark = ",", digits =0) %>% 
  formatCurrency('total_miles',currency = "", interval = 3, mark = ",", digits =0) %>% 
  formatCurrency('total_area_miles2',currency = "", interval = 3, mark = ",", digits =0) %>% 
  formatCurrency('divides',currency = "", interval = 3, mark = ",", digits =0) %>% 
  formatCurrency('nexus',currency = "", interval = 3, mark = ",", digits =0) %>% 
  formatCurrency('hydrolocations',currency = "", interval = 3, mark = ",", digits =0) %>% 
  formatStyle(0, target = "row", fontWeight = styleEqual(dim(m)[1], "bold"))

  
DT
```

