---
title: "Sprint 0: Experimenting with representation"
author: "Mike Johnson"
date: "`r Sys.Date()`"
output: html_document
out_dir: docs
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Dependencies
library(dplyr)
library(units)
library(sf)
library(jsonlite)
```


# Problem/Background

Three tenets of `hy_features` are:

- a catchment realization must perform the functions of a *flowpath* and a *catchment* area.
- Any catchment realization may be composed of a network of other catchment realizations.
- Catchment and nexus provide identity and topology but do _not_ have attributes or geometry.

- But, we *need* geometry to inform processing, and we need that geometry to associate flow paths and catchment intercahangbly.

We have 2 distinct files with a shared ID (comid):
```{r}
source('../R/prep_flownetwork.R')

og_cat <- read_sf("../data/huc1805000502-catchment.gpkg") %>% 
   dplyr::select(comid = FEATUREID)

og_fl = read_sf("../data/huc1805000502-flowlines.gpkg") %>% 
  prep_flownetwork() 
```

Lets create a single object that contains both "realizations" is a single go:

```{r}
source("../R/toy_functions.R")

hy = make_hy(og_fl, og_cat)
head(hy)
```

Looking at the header of the print, we see that the 'active' geometry column is called `cat_geom`.
Thus, if we plot, we get:

```{r}
plot(st_geometry(hy))
```

But, often we want/need the flowpath represneation! So lets "activate" the flow path:

```{r}
hy = activate(hy, "fl")
plot(st_geometry(hy))
```

Easy plot function:

```{r}
plot_hy(hy)
```

## Modification
Awesome! Now what about "realization" centric calcuations like length/area?

```{r}
realize(hy, "fl") %>% st_length()
realize(hy, "fl") %>% st_area()
realize(hy, "cat") %>% st_area()
```

Flow path should be LINESTRINGS, not MULTILINESTRINGS. Always!

```{r}
hy = multi_to_line(realize(hy, "fl"))
```

## {Graph}ing

We can also look at "spatial graph" representations of this `hy` object. 
Here we enforce, a topologic order as well:

```{r}
hy_net = make_graph(hy)
hy_net %>% tidygraph::activate("edges"),"cat")[[edges]]
plot(realize(activate(hy_net, "edges"), "cat"),add = TRUE)
```
