---
title: "Hydrofabric Data Model and Alignment"
author: "Mike Johnson"
date: "2022-10-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(hydrofab)
library(dm)
library(DBI)
library(sf)
```

```{r, echo = FALSE}
hf = list(
lookup_table = data.frame(
  NHDPlusV2_COMID = numeric(1L),
  NHDPlusV2_COMID_part = integer(1L),
  id = integer(1L),
  divide_id = integer(1L),
  mainstem = integer(1L),
  poi_id = integer(1L),
  poi_type = integer(1L),
  poi_value = integer(1L)
),

reference_flowlines = data.frame(
  id = integer(1L),
  toid = integer(1L),
  member_comid = character(1L),
  mainstem = integer(1L),
  lengthkm = numeric(1L),
  tot_drainage_areasqkm = numeric(1L),
  slope = numeric(1L),
  order = numeric(1L),
  hydroseq = numeric(1L),
  areasqkm = numeric(1L),
  gnis_id = numeric(1L),
  gnis_name = character(1L),
  `wbareacomi_` = integer(1L),
  poi_id = integer(1L),
  divide_id = integer(1L),
  geometry = numeric(1L)
),


mapped_POIs = data.frame(
  poi_id = integer(1L),
  id     = integer(1L),
  type_HUC12 = character(1L),
  type_gages = character(1L),
  type_te    = character(1L),
  type_nid   = character(1L),
  type_wbin  = character(1L),
  type_wbout = character(1L),
  type_conf  = character(1L),
  type_term  = character(1L),
  type_elev  = character(1L),
  type_travel = character(1L),
  type_con    = character(1L),
  nexus = character(1L),
  snapped = numeric(1L),
  geometry = numeric(1L)
),

reference_catchments = data.frame(
  divide_id = integer(1L),
  toid = integer(1L),
  areasqkm = numeric(1L),
  type = character(1L),
  geometry = numeric(1L)
),

reference_network = data.frame(
  id = integer(1L),
  toid = integer(1L),
  divide_id = integer(1L),
  lengthkm = numeric(1L),
  areasqkm = numeric(1L),
  tot_drainage_areasqkm = numeric(1L),
  `wbareacomi_` = integer(1L),
  hydroseq = numeric(1L),
  mainstem = numeric(1L)
),

`WB_` = data.frame(
  `wbareacomi_` = integer(1L),
  gnis_id = numeric(1L),
  gnis_name = character(1L),
  wb_areasqkm = numeric(1L),
  wb_elevation = numeric(1L),
  wb_mean_depth = numeric(1L),
  wb_lake_volume = numeric(1L),
  wb_max_depth = numeric(1L),
  wb_source = numeric(1L),
  geometry = numeric(1L)
)
)




meta = tibble::tribble(
  ~Attribute, ~Description,
"NHDPlusV2_COMID",          "The NHDPlusCOMID",
"NHDPlusV2_COMID_part",     "If the original NHDPlus COMID was split, the subpart.	Parts increase from outlet to inlet",
"id",                       "A hydrofabric specfic, globaly unique flowpath/flowline identifier",
"divide_id",                "A hydrofabric specfic, globaly unique divide identifier",
"mainstem",                 "the primary downstream segment of a river, as contrasted to its tributaries",
"poi_id",                   "A hydrofabric specifc, globaly unique Point of Interest identifier",
"poi_type",                 "The Point of Interest type",
"poi_value",                "The origional identifier of the POI in the native dataset",
"toid",                     "The identifier of the directly downstream flowpath/flowline",
"lengthkm",                 "The length in kilometers of the flowpath element",
"tot_drainage_areasqkm",    "The total upstream area contributing to the flowpath/flowline",
"order",                    "Strahler stream order",
"hydroseq",                 "VPU based hydrologic sort. Increases from downstream to upstream",
"areasqkm",                 "The area of the incremental divide for a given flowpath/flowline",
"gnis_id",                  "Geographic Names Information System Identifier",
"gnis_name",                "Geographic Names Information System Name",
"wbareacomi_",              "NHDPlusV2 Water body Common Identifier",
"geometry",                 "Simple Features Geometry",
"type_HUC12",               "Point of Interest made at a HUC12 outlet (contains WBD HUC12 ID)",
"type_gages",               "Point of Interest made at an NWIS gage (contains eight digit NWIS ID)",
"type_te",                  "Point of Interest made at a Thermoelectic Plant (contains native dataset ID)",
"type_nid",                 "Point of Interest made at an Army Corp Dam (contains NID ID)",
"type_wbin",                "Point of Interest made at a Waterbody Inflow (contains NHDPlusV2 waterbody ID)",
"type_wbout",               "Point of Interest made at a Waterbody Outflow (contains NHDPlusV2 waterbody ID)",
"type_conf",                "Point of Interest made at a network confluence (contains NHDPlusV2 waterbody ID)",
"type_term",                "Point of Interest made at a network terminal (contains NHDPlusV2 COMID)",
"type_elev",                "Point of Interest made at a network elevation break (contains NHDPlusV2 COMID)",
"type_travel",              "Point of Interest made at a network travel time break (contains NHDPlusV2 COMID)",
"type",                     "Type of divide (network, internal, coastal)",
"wb_areasqkm",              "Waterbody area",
"wb_elevation",             "Waterbody elevation",
"wb_mean_depth",            "Waterbody mean depth",
"wb_lake_volume",           "Waterbody lake volume",
"wb_max_depth",             "Waterbody max depth",
"wb_source",                "Waterbody source"
)


dm = dm::dm(lookup_table = hf$lookup_table,
            flowlines = hf$reference_flowlines,
            mapped_POIs = hf$mapped_POIs,
            divides = hf$reference_catchments,
            network = hf$reference_network,
            WB_ = hf$`WB_`) %>% 

dm_add_pk(flowlines, id)  %>% 
  dm_add_pk(mapped_POIs, poi_id)  %>% 
  dm_add_pk(WB_, `wbareacomi_`)  %>% 
  dm_add_pk(network, id)  %>% 
  dm_add_pk(divides, divide_id) %>% 
  dm_add_pk(lookup_table, id) %>% 
  
  dm_add_fk(flowlines,  `wbareacomi_`, WB_) %>% 
  dm_add_fk(network,  `wbareacomi_`, WB_) %>% 
  dm_add_fk(flowlines,  poi_id, mapped_POIs) %>% 
  dm_add_fk(flowlines,  divide_id, divides) %>% 
  
  dm_add_fk(mapped_POIs,  id, network) %>% 
  dm_add_fk(lookup_table,  id, network) %>% 
  dm_add_fk(lookup_table,  id, flowlines) %>% 
  dm_add_fk(flowlines,  id, network) %>% 
  dm_add_fk(lookup_table,  divide_id, divides) %>% 
  dm_add_fk(lookup_table,  poi_id, mapped_POIs) %>% 
  dm_set_colors(red = flowlines, 
                orange = divides,
                green = lookup_table, 
                yellow = mapped_POIs,
                purple = WB_, 
                violet = network)


df = data.frame(names = names(unlist(hf))) %>% 
             tidyr::separate(names, sep = "[.]", into  = c('layer', 'Attribute')) %>% 
  group_by(Attribute) %>% 
  summarise(layers = paste(layer, collapse = ", "))


meta = left_join(meta, df, by = "Attribute")

```

# Current State of Affairs {.tabset}

## Proposed Idealized Data Model (10/14/2022)

```{r, echo = FALSE, fig.height=8, fig.width=8, fig.show='hold'}
dm_draw(dm,  view_type = "all", column_types = TRUE)
```

```{r, echo = FALSE}
DT::datatable(meta, options = list(pageLength = 35, dom = 't'), rownames = FALSE)
```


## Reference

### Exisiting Reference Fabric

```{r, echo = FALSE, fig.height=8, fig.width=8, fig.show='hold'}
ref = get_hydrofabric(VPU = "01", "reference", dir = '/Volumes/Transcend/ngen/CONUS-hydrofabric/')

db <- dbConnect(RSQLite::SQLite(), ref)

OG = dm_from_con(db, learn_keys = FALSE)[c(st_layers(ref)$name)] %>% 
  dm_select(reference_flowline, -fid) %>% 
  dm_select(POIs, -fid) %>% 
  dm_select(WB_01, -fid) %>% 
  dm_select(reference_network, -fid) %>% 
  dm_select(reference_catchment, -fid) %>% 
  dm_select(lookup_table, -fid)  %>% 
  dm_set_colors(red = reference_flowline, 
                orange = reference_catchment,
                green = lookup_table, 
                yellow = POIs,
                purple = WB_01, 
                violet = reference_network,
  )

dm_draw(OG, view_type = "all", rank = "RL", column_types = TRUE) 

```


### Idealized Reference Fabric

```{r, echo = FALSE, fig.height=8, fig.width=8, fig.show='hold'}
dm_draw(dm, view_type = "all", rank = "RL", column_types = TRUE)
```

### Required Changes: 

#### 1. WB

- [ ] Change table name from `WB_*` to `WB`
- [ ] Change all names to lower case
 
- [ ] *Remove*:
    - [ ]  `COMID`
    - [ ]  `FDATE`
    - [ ]  `RESOLUTION`
    - [ ]  `FTYPE`
    - [ ]  `FCODE`
    - [ ]  `Shape_Length`
    - [ ]  `Shape_Area`
    - [ ]  `ONOFFNET`
    - [ ]  `PurpCode`
    - [ ]  `PurpDesc`
    
- [ ] *Add*: 
    - [ ] NA
    
- [ ] *Change*:
    - [ ] NA
    
#### 2. POIs

- [ ]  Change all names to lower case
 
- [ ] *Remove*:
   - [ ]  NA
   
- [ ] *Add*: 
   - [ ]  NA
   
- [ ] *Change*:
    - [ ] `Identifier` --> `poi_id`
    - [ ] `COMID` --> `id`
    
#### 3. Lookup Table

- [ ]  Change all names to lower case
 
- [ ] *Remove*:
    - [ ]  NA
   
- [ ] *Add*: 
    - [ ] `NHDPlusV2_COMID_part` (always 1)
    - [ ] `id` (equals COMID in reference)
    
- [ ] *Change*:
    - [ ] `value` --> `poi_value`
    - [ ] `type` --> `poi_type`
    - [ ] `realizied_catchmentID` --> `divide_id`
    
#### 4. Reference network

- [ ] Change all names to lower case
 
- [ ] *Remove* (all of these can be looked up by id/comid):
    - [ ] `terminalpa`
    - [ ] `dnlevelpat`
    - [ ] `dnhydroseq`
    - [ ] `reachcode`
    - [ ] `frommeas`
    - [ ] `tomeas`
    - [ ] `pathlength`
    - [ ] `arbolatesu`
    - [ ] `ftype`
    - [ ] `fcode`
    - [ ] `vpuid`
    - [ ] `rpuid` 
   
- [ ] *Add*: 
    - [ ] `divide_id`
   
- [ ] *Change*:
    - [ ] `comid` --> `id`
    - [ ] `tocomid` --> `toid`
    - [ ] `totdasqkm` --> `tot_draiange_areasqkm`
    - [ ] `levelpathi` -->  `mainstem`
    
#### 5. Reference catchments

- [ ] Change all names to lower case
- [ ] Change table name to `reference_divides`
 
 - *Remove*:
    - [ ] `VPUID`
    - [ ] `RPUID`
    - [ ] `hy_cats`
    - [ ] `full_cats`
   
- [ ] *Add*: 
    - [ ] `type`
   
- [ ] *Change*:
    - [ ] `FEATUREID` --> `divide_id`
    - [ ] `tocomid` --> `toid`
    - [ ] `totdasqkm` --> `tot_draiange_areasqkm`
    - [ ] `levelpathi` -->  `mainstem`
    
#### 6. Reference Flowlines

- [ ] Change all names to lower case
 
- [ ] *Remove*:
    - [ ] `FromNode`
    - [ ] `ToNode`
    - [ ] `StartFlag`
    - [ ] `StreamCalc`
    - [ ] `Diveregence`
    - [ ] `DnMinorHyd`
    - [ ] `FCODE`
    - [ ] `REACHCODE`
    - [ ] `FromMeas`
    - [ ] `ToMeas`
    - [ ] `ArbolateSu`
    - [ ] `TernimalPa`
    - [ ] `Pathlength`
    - [ ] `DnLevelPat`
    - [ ] `DnHydroseq`
    - [ ] `TerminalFl`
    - [ ] `streamleve`
    - [ ] `vpuin`
    - [ ] `vpuout`
    - [ ] `wbareaype`
    - [ ] `slopelenkm`
    - [ ] `FTYPE`
    - [ ] `hwnodesqkm`
    - [ ] `RPUID`
    - [ ] `VPUID`
    - [ ] `roughness`
    - [ ] `dend`
    - [ ] `WB`
    - [ ] `minNext`
    - [ ] `poi`
    - [ ] `struct_POI`
    - [ ] `struct_net`
   
- [ ] *Add*: 
    - [ ] `member_comid`
   
- [ ] *Change*:
    - [ ] `COMID` --> `id`
    - [ ] `toCOMID` --> `toid`
    - [ ] `TotDASqKM` --> `tot_draiange_areasqkm`
    - [ ] `LevelPathI` -->  `mainstem`
    - [ ] `StreamOrd` --> order
    
## Refactor

### Exisiting Refactor Fabric

```{r, echo = FALSE, fig.height=8, fig.width=8, fig.show='hold'}
rf = get_hydrofabric(VPU = "01", "refactor", dir = '/Volumes/Transcend/ngen/CONUS-hydrofabric/')

db <- dbConnect(RSQLite::SQLite(), rf)

OG = dm_from_con(db, learn_keys = FALSE)[c(st_layers(rf)$name)] %>% 
  dm_select(refactored_flowpaths, -fid) %>% 
  dm_select(mapped_POIs, -fid) %>% 
  dm_select(catchment_network, -fid) %>% 
  dm_select(refactored_divides, -fid) %>% 
  dm_select(lookup_table, -fid)  %>% 
  dm_set_colors(red = refactored_flowpaths, 
                orange = refactored_divides,
                green = lookup_table, 
                yellow = mapped_POIs,
                violet = catchment_network
  )


dm_draw(OG, view_type = "all", rank = "RL", column_types = TRUE) 

```

### Idealized Refactored Fabric

```{r, echo = FALSE, fig.height=8, fig.width=8, fig.show='hold'}
dm = dm::dm(lookup_table = hf$lookup_table,
            flowpaths = mutate(hf$reference_flowlines, wbareacomi_ = NULL),
            mapped_POIs = hf$mapped_POIs,
            divides = hf$reference_catchments,
            network = mutate(hf$reference_network, wbareacomi_ = NULL)) %>% 

dm_add_pk(flowpaths, id)  %>% 
  dm_add_pk(mapped_POIs, poi_id)  %>% 
  dm_add_pk(network, id)  %>% 
  dm_add_pk(divides, divide_id) %>% 
  dm_add_pk(lookup_table, id) %>% 
  
  dm_add_fk(flowpaths,  poi_id, mapped_POIs) %>% 
  dm_add_fk(flowpaths,  divide_id, divides) %>% 
  
  dm_add_fk(mapped_POIs,  id, network) %>% 
  dm_add_fk(lookup_table,  id, network) %>% 
  dm_add_fk(lookup_table,  id, flowpaths) %>% 
  dm_add_fk(flowpaths,  id, network) %>% 
  dm_add_fk(lookup_table,  divide_id, divides) %>% 
  dm_add_fk(lookup_table,  poi_id, mapped_POIs) %>% 
  dm_set_colors(red = flowpaths, 
                orange = divides,
                green = lookup_table, 
                yellow = mapped_POIs,
                violet = network)


#dm_draw(dm,  rank = "RL", column_types = TRUE)


dm_draw(dm, view_type = "all", rank = "RL", column_types = TRUE)

```

### Required Changes: 


#### 1. mapped_POIs

 - Change all names to lower case
 
 - *Remove*:
    - `toid`, `set`, `comid`, `member_COMID`, `TotDASqKM`, `DnHydroseq`,`type`, `member_COMID`
   
 - *Add*: 
    - NA
    
 - *Change*:
    - `identifier` --> `poi_id`
    - `geom` --> `geometry`
    
#### 2. Lookup Table

 - Change all names to lower case
 
 - *Remove*:
    - member_COMID
   
   
 - *Add*: 
    - `NHDPlusV2_COMID_part`
    - `divide_id`
    - `POI_ID`
    - `POI_TYPE`
    - `POI_VALUE`
    
    
 - *Change*:
    - `reconciled_ID` --> `id`
    - `LevelPathID` --> `mainstem`
    
#### 3.  Network

- Change all names to lower case
 
 - *Remove*:
   - NA
   
 - *Add*: 
    - `hydroseq`, `tot_drainage_areasqkm`, `divide_id`
   
 - *Change*:
    - `levelpathid` --> `mainstem`
    
#### 4. divides 

- Change all names to lower case
- change table name to `reference_divides`
 
 - *Remove*:
   - `rpu`, `member_COMID`
   
 - *Add*: 
  - `type`, `toid`
   
 - *Change*:
    - `ID` --> `divide_id`
    
#### 5. Reference Flowlines

- Change all names to lower case
 
 - *Remove*:
  - `refactor_id`
   
   
 - *Add*: 
    - gnis_id, gnis_name, hydroseq order, slope, areasqkm, poi_id, divide_id
   
 - *Change*:
    - `TotDASqKM` --> `tot_draiange_areasqkm`
    - `LevelPathID` -->  `mainstem`
    
## Minimal

### Exisiting Minimal Fabric

```{r, echo = FALSE, fig.height=8, fig.width=8, fig.show='hold'}
rm = get_hydrofabric(VPU = "01", "minimal", dir = '/Volumes/Transcend/ngen/CONUS-hydrofabric/', overwrite = FALSE)

db <- dbConnect(RSQLite::SQLite(), rm)

OG = dm_from_con(db, learn_keys = FALSE)[c(st_layers(rm)$name)] %>% 
  dm_select(aggregated_flowpaths, -fid) %>% 
  dm_select(mapped_POIs, -fid) %>% 
  dm_select(catchment_network, -fid) %>% 
  dm_select(aggregated_divides, -fid) %>% 
  dm_select(lookup_table, -fid)  %>% 
  dm_set_colors(red = aggregated_flowpaths, 
                orange = aggregated_divides,
                green = lookup_table, 
                yellow = mapped_POIs,
                violet = catchment_network
  )

dm_draw(OG, view_type = "all", rank = "RL", column_types = TRUE) 
```

### Idealized Minimal Fabric

```{r, echo = FALSE, fig.height=8, fig.width=8, fig.show='hold'}
dm = dm::dm(lookup_table = hf$lookup_table,
            flowpaths = mutate(hf$reference_flowlines, wbareacomi_ = NULL),
            mapped_POIs = hf$mapped_POIs,
            divides = hf$reference_catchments,
            network = mutate(hf$reference_network, wbareacomi_ = NULL)) %>% 

dm_add_pk(flowpaths, id)  %>% 
  dm_add_pk(mapped_POIs, poi_id)  %>% 
  dm_add_pk(network, id)  %>% 
  dm_add_pk(divides, divide_id) %>% 
  dm_add_pk(lookup_table, id) %>% 
  
  dm_add_fk(flowpaths,  poi_id, mapped_POIs) %>% 
  dm_add_fk(flowpaths,  divide_id, divides) %>% 
  
  dm_add_fk(mapped_POIs,  id, network) %>% 
  dm_add_fk(lookup_table,  id, network) %>% 
  dm_add_fk(lookup_table,  id, flowpaths) %>% 
  dm_add_fk(flowpaths,  id, network) %>% 
  dm_add_fk(lookup_table,  divide_id, divides) %>% 
  dm_add_fk(lookup_table,  poi_id, mapped_POIs) %>% 
  dm_set_colors(red = flowpaths, 
                orange = divides,
                green = lookup_table, 
                yellow = mapped_POIs,
                violet = network)


dm_draw(dm, view_type = "all", rank = "RL", column_types = TRUE)

```

## Uniform

### Exisiting Uniform Fabric

```{r, echo = FALSE, fig.height=8, fig.width=8, fig.show='hold'}
rm = '/Volumes/Transcend/ngen/CONUS-hydrofabric/pre-release/global_uniform_01.gpkg'

db <- dbConnect(RSQLite::SQLite(), rm)

OG = dm_from_con(db, learn_keys = FALSE)[c(st_layers(rm)$name)] %>% 
  dm_select(flowpaths, -fid) %>% 
  dm_select(mapped_POIs, -fid) %>% 
  dm_select(flowpath_edge_list, -fid) %>% 
  dm_select(divides, -fid) %>% 
  dm_select(lookup_table, -fid)  %>% 
  dm_set_colors(red = flowpaths, 
                orange = divides,
                green = lookup_table, 
                yellow = mapped_POIs,
                violet = flowpath_edge_list
  )

dm_draw(OG, view_type = "all", rank = "RL", column_types = TRUE) 
```

### Idealized Minimal Fabric
